\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue
}

\title{Mathematical Theory Compendium}
\author{Srikrishnaa Vadivel}
\date{}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\g}{\mathfrak{g}}
\newcommand{\dd}{\,\mathrm{d}}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\pdd}[2]{\frac{\partial^2 #1}{\partial #2^2}}

\begin{document}
\maketitle
\tableofcontents

\section{Overview}

This note is an attempt to write broad areas of mathematics through compact and
reusable equations. The emphasis is not on encyclopedic completeness. The goal
is to show a style of mathematical organization: define the objects cleanly,
state the structural laws precisely, and then express higher theories as
controlled extensions of the same notation.

The three main threads here are:
\begin{itemize}
  \item algebra as the study of operations and the laws they satisfy,
  \item analysis as the study of functions, variation, and differential
  equations,
  \item geometry as the study of smooth spaces, tensors, and curvature.
\end{itemize}

\section{Algebra}

\subsection{Terms and Operations}

Let
\[
  x_i \in X
\]
be symbols in a set \(X\). Let
\[
  f_i : X^{n_i} \to X
\]
be an operation of arity \(n_i\). In prefix notation,
\[
  f_i x_1 \cdots x_{n_i} = f_i(x_1,\dots,x_{n_i}).
\]

This viewpoint is useful because it makes the algebraic laws look like
equational templates. A structure is then determined by the operations it
admits together with the identities they satisfy.

\subsection{Fundamental Algebraic Laws}

\paragraph{Closure.}
\[
  f(x,y) \in X.
\]

\paragraph{Commutativity.}
\[
  f(x,y) = f(y,x).
\]

\paragraph{Associativity.}
\[
  f(f(x,y),z) = f(x,f(y,z)).
\]

\paragraph{Identity.}
There exists \(e \in X\) such that
\[
  f(e,x) = x,
  \qquad
  f(x,e) = x.
\]

\paragraph{Inverse.}
For every \(x \in X\), there exists \(x^{-1}\) such that
\[
  f(x,x^{-1}) = e.
\]

\paragraph{Idempotence.}
\[
  f(x,x) = x.
\]

\paragraph{Absorption.}
For two operations \(f\) and \(g\),
\[
  f(x,g(x,y)) = x.
\]

\paragraph{Distributivity.}
For two operations \(f\) and \(g\),
\[
  f(x,g(y,z)) = g(f(x,y),f(x,z)).
\]

\subsection{Basic Algebraic Structures}

\paragraph{Magma.}
A magma is a set \(X\) equipped with one binary operation
\[
  f : X \times X \to X.
\]

\paragraph{Semigroup.}
A semigroup is a magma satisfying associativity.

\paragraph{Monoid.}
A monoid is a semigroup with an identity element.

\paragraph{Group.}
A group is a monoid in which every element has an inverse.

\paragraph{Abelian group.}
An Abelian group is a group that is also commutative.

These structures form a natural ladder: each new notion is obtained by adding
one more law to the previous one.

\subsection{Rings and Fields}

A ring \(R\) has two operations \(+\) and \(\cdot\) such that
\[
  (R,+)
\]
is an Abelian group,
\[
  (R,\cdot)
\]
is a monoid, and multiplication distributes over addition:
\[
  a(b+c) = ab + ac,
  \qquad
  (a+b)c = ac + bc.
\]

A field \(\F\) is a commutative ring for which
\[
  (\F \setminus \{0\}, \cdot)
\]
is an Abelian group. In practice, the field condition is what allows division by
nonzero scalars and therefore makes linear theory behave cleanly.

\subsection{Modules and Vector Spaces}

Let \(R\) be a ring and \(M\) an Abelian group. A module structure is a scalar
multiplication
\[
  R \times M \to M,
  \qquad
  (r,v) \mapsto rv,
\]
satisfying
\[
  r(u+v) = ru + rv,
\]
\[
  (r+s)v = rv + sv,
\]
\[
  (rs)v = r(sv).
\]

When the base ring is a field \(\F\), the module is a vector space. This is the
point where algebra becomes linear algebra: vector addition and scalar
multiplication are governed by the same distributive pattern, but now over a
field rather than an arbitrary ring.

\subsection{Associative and Lie Algebras}

An associative algebra over a field \(\F\) is a vector space \(A\) equipped with
a bilinear multiplication
\[
  A \times A \to A
\]
such that
\[
  (xy)z = x(yz).
\]

A Lie algebra is a vector space \(\g\) with a bilinear bracket
\[
  [-,-] : \g \times \g \to \g
\]
satisfying
\[
  [x,y] = -[y,x],
\]
and the Jacobi identity
\[
  [x,[y,z]] + [y,[z,x]] + [z,[x,y]] = 0.
\]

The contrast is important: associative algebras encode composition-like
products, while Lie algebras isolate infinitesimal commutation structure.

\subsection{Constructed Algebras}

\paragraph{Tensor algebra.}
For a vector space \(V\),
\[
  T(V) = \bigoplus_{k=0}^{\infty} V^{\otimes k}.
\]

\paragraph{Exterior algebra.}
The exterior algebra imposes antisymmetry through the relation
\[
  v \wedge v = 0.
\]

\paragraph{Clifford algebra.}
For a quadratic form \(g\),
\[
  uv + vu = 2g(u,v).
\]

These constructions are useful because they move from a bare vector space to
larger algebras that remember multilinear, antisymmetric, and metric structure.

\subsection{Normed and Involutive Algebras}

A Banach algebra is an algebra \(A\) with a norm \(\|\cdot\|\) such that \(A\)
is complete and
\[
  \|xy\| \le \|x\| \|y\|.
\]

A \(C^*\)-algebra is a Banach algebra equipped with an involution
\[
  * : A \to A
\]
satisfying
\[
  (ab)^* = b^* a^*,
\]
\[
  (a^*)^* = a,
\]
\[
  \|a^* a\| = \|a\|^2.
\]

\section{Analysis}

\subsection{Functions and Composition}

A function is a map
\[
  f : X \to Y.
\]
Composition is defined by
\[
  (f \circ g)(x) = f(g(x)).
\]

The main point is that analysis starts with maps, but quickly moves to
structured maps: differentiable maps, integral operators, variational
functionals, and solution operators for differential equations.

\subsection{Multi-Index Notation}

Let
\[
  J = (j_1,\dots,j_n) \in \N^n,
  \qquad
  |J| = \sum_{i=1}^n j_i.
\]
For a sufficiently smooth function \(u\),
\[
  D_J u =
  \frac{\partial^{|J|} u}
  {\partial x_1^{j_1} \cdots \partial x_n^{j_n}}.
\]

This notation compresses higher derivatives into a form that scales well when
moving from ordinary calculus to PDEs and variational systems.

\subsection{Jets and Total Derivatives}

Let \(u^\alpha(x)\) be unknown fields. The associated jet coordinates are
\[
  (x^i, u^\alpha, u^\alpha_J).
\]

The total derivative in the \(x^i\)-direction is
\[
  D_i
  =
  \frac{\partial}{\partial x^i}
  +
  \sum_J
  u^\alpha_{J,i}
  \frac{\partial}{\partial u^\alpha_J}.
\]

This is the right derivative when the variables include not only the fields but
also all of their derivatives.

\subsection{Integration and Functionals}

For a domain \(\Omega \subset \R^n\),
\[
  \int_\Omega f(x)\,\dd x
\]
is the integral of a scalar density over \(\Omega\).

A variational functional has the form
\[
  \mathcal{F}[u]
  =
  \int_\Omega
  L(x,u,u_J)\,\dd x,
\]
where \(L\) is the Lagrangian density. This turns a field \(u\) into a scalar
quantity and is the basic setup behind the calculus of variations.

\subsection{Euler--Lagrange Equation}

The stationarity condition for \(\mathcal{F}[u]\) is
\[
  \sum_J
  (-D)_J
  \frac{\partial L}{\partial u_J}
  = 0.
\]

This equation is one of the central bridges between analysis and physics: it
converts an extremization principle into a differential equation.

\subsection{Differential Equations and Boundary Conditions}

A general nonlinear differential equation can be written as
\[
  F(x,u,u_J) = 0.
\]

Standard boundary conditions include:

\paragraph{Dirichlet.}
\[
  u = g.
\]

\paragraph{Neumann.}
\[
  \pd{u}{n} = g.
\]

\paragraph{Robin.}
\[
  au + b \pd{u}{n} = g.
\]

\subsection{Weak Forms}

Instead of solving a differential equation pointwise, one often seeks
\[
  u \in V
\]
such that
\[
  a(u,v) = \ell(v)
  \qquad
  \text{for all } v \in V.
\]

This formulation is central in finite elements, functional analysis, and modern
PDE theory because it weakens differentiability requirements while preserving
the core operator structure.

\subsection{Banach and Hilbert Spaces}

A Banach space is a complete normed vector space with norm
\[
  \|u\|.
\]

A Hilbert space is a complete inner-product space with inner product
\[
  \langle u,v \rangle.
\]

The extra inner-product structure in Hilbert spaces gives orthogonality,
projection, and spectral theory in a particularly clean form.

\subsection{Operators and Spectral Problems}

A linear operator is a map
\[
  L : V \to W.
\]

The corresponding eigenvalue problem is
\[
  Lu = \lambda u.
\]

This equation organizes much of analysis around invariant modes, frequencies,
and stability behavior.

\section{Geometry}

\subsection{Smooth Manifolds}

A smooth manifold is a topological space \(M\) equipped with an atlas
\[
  \{(U_i,\phi_i)\},
\]
where
\[
  \phi_i : U_i \to \R^n
\]
are compatible coordinate charts.

The point is that geometry is built by moving between local Euclidean
coordinates and globally defined objects that do not depend on a particular
chart.

\subsection{Tangent, Cotangent, and Tensor Spaces}

At each point \(p \in M\), the tangent space is
\[
  T_p M,
\]
and the cotangent space is
\[
  T_p^* M.
\]

From these, one builds tensor bundles
\[
  T^r_s M.
\]

These are the natural homes for vector fields, covectors, metrics, curvature,
and all other multilinear geometric quantities.

\subsection{Differential Forms and Exterior Calculus}

A differential \(k\)-form is an element
\[
  \omega \in \Omega^k(M).
\]

The exterior derivative is
\[
  d : \Omega^k(M) \to \Omega^{k+1}(M),
  \qquad
  d^2 = 0.
\]

The exterior product is
\[
  \omega \wedge \eta.
\]

This package of ideas turns geometry into a calculus of antisymmetric tensors
that behaves well under coordinate change.

\subsection{Metrics and Connections}

A metric is a symmetric \((0,2)\)-tensor
\[
  g = g_{ij}\,\dd x^i \otimes \dd x^j.
\]

A connection differentiates tensor fields. On vector fields, one writes
\[
  \nabla_X Y,
\]
or locally
\[
  \nabla_{\partial_i} \partial_j
  =
  \Gamma^k_{ij}\,\partial_k.
\]

Connections encode how vectors are transported and compared across nearby
points of a manifold.

\subsection{Curvature}

The curvature operator is
\[
  R(X,Y)Z.
\]

In coordinates, its components are written
\[
  R^i{}_{jkl}.
\]

The Ricci tensor is obtained by contraction:
\[
  R_{ij}.
\]

The scalar curvature is a second contraction:
\[
  R = g^{ij} R_{ij}.
\]

Together these quantities measure how far the geometry is from being flat.

\subsection{Geodesics}

A geodesic \(\gamma(t)\) satisfies
\[
  \frac{D \dot{\gamma}}{\dd t} = 0.
\]

This is the intrinsic version of straight-line motion: the acceleration
vanishes with respect to the connection, not necessarily in ambient Euclidean
coordinates.

\subsection{Bundles and Gauge Geometry}

A fiber bundle has the form
\[
  F \hookrightarrow E \to M.
\]

A principal \(G\)-bundle is written
\[
  G \hookrightarrow P \to M.
\]

A connection one-form is denoted by
\[
  A.
\]

Its curvature two-form is
\[
  F = \dd A + A \wedge A.
\]

The covariant exterior derivative is
\[
  D = \dd + A.
\]

This notation is the natural language for gauge theory, differential geometry,
and many geometric formulations of physics.

\subsection{Hodge Theory and Integral Identities}

The Hodge star operator is
\[
  * : \Omega^k(M) \to \Omega^{n-k}(M).
\]

The Laplace--Beltrami operator is
\[
  \Delta = \dd \delta + \delta \dd.
\]

Stokes' theorem closes the circle between local differentiation and global
integration:
\[
  \int_M \dd \omega = \int_{\partial M} \omega.
\]

\section{Closing View}

One way to read the document is as a progression in which each subject refines
the last:
\begin{itemize}
  \item algebra studies operations and the identities they satisfy,
  \item analysis studies functions, operators, and variational equations built
  from those operations,
  \item geometry studies spaces on which those operators and equations become
  coordinate-invariant.
\end{itemize}

The long-term aim is not merely to collect formulas, but to express a style of
thinking: start from the primitive objects, choose notation that scales, and let
the equations reveal the structure.

\end{document}
